Ecosystem carbon balance in the Hawaiian Islands under different scenarios of future climate and land use change

### 2021-01-15 Manuscript/HI_Model_ms.Rmd
updated manuscript files

The State of Hawaiʻi exemplifies the challenges of projecting future interactive effects of climate and land use change on ecosystem carbon balance. The main Hawaiian Islands are a complex mosaic of natural and human-dominated landscapes overlain by steep climate gradients across relatively short distances. Mean annual temperature ranges from ~4-24° C [@giambelluca_evapotranspiration_2014] and mean annual rainfall ranges from ~180-9500 mm [@giambelluca_online_2013], with minimal seasonal variation in climate. Temperatures have risen rapidly in the Hawaiian Islands since the mid 1970's [@giambelluca_secular_2008] and a long-term drying trend has persisted since the early 1920s [@frazier_spatial_2017], resulting in reduced forest biomass and productivity [@barbosa_effects_2017]. These same drying and warming trends have increased the frequency and intensity of wildland fire [@trauernicht_contemporary_2015; @trauernicht_vegetationrainfall_2019] with predictable negative effects on ecosystem carbon balance [@selmants_baseline_2017]. Ecosystem carbon stocks across the main Hawaiian Islands have also been strongly influenced by the legacy of past land-use change [@asner_high-resolution_2011]. Thousands of hectares of land were deforested beginning in the late 19th century to clear land for sugar plantations and cattle pasture [@cuddihy_alteration_1990]. Beginning in the latter half of the 20th century, this agricultural land has been steadily converted to urban areas, commercial forestry plantations, or simply abandoned and colonized by non-native grass species [@suryanata_diversified_2009; @perroy_evolving_2016].

>>>>

The State of Hawaiʻi exemplifies the challenges of projecting future interactive effects of climate and land use change on ecosystem carbon balance. The main Hawaiian Islands are a complex mosaic of natural and human-dominated landscapes overlain by steep climate gradients across relatively short distances. Mean annual temperature ranges from ~4-24° C [@giambelluca_evapotranspiration_2014] and mean annual rainfall ranges from ~180-9500 mm [@giambelluca_online_2013], with minimal seasonal variation in climate. Temperatures have risen rapidly in the Hawaiian Islands since the mid 1970's [@giambelluca_secular_2008] and a long-term drying trend has persisted since the early 1920s [@frazier_spatial_2017], resulting in reduced forest biomass and productivity [@barbosa_effects_2017]. These same drying and warming trends have increased the frequency and intensity of wildland fire [@trauernicht_contemporary_2015; @trauernicht_vegetationrainfall_2019] with predictable negative effects on ecosystem carbon balance [@selmants_baseline_2017]. Ecosystem carbon stocks across the main Hawaiian Islands have also been strongly influenced by the legacy of past land-use change [@osher_effect_2003; @asner_high-resolution_2011]. Thousands of hectares of land were deforested beginning in the late 19th century to clear land for sugar plantations and cattle pasture [@cuddihy_alteration_1990]. Beginning in the latter half of the 20th century, this agricultural land has been steadily converted to urban areas, commercial forestry plantations, or simply abandoned and colonized by non-native grass species [@suryanata_diversified_2009; @perroy_evolving_2016].
